\large{\bf Large effects and the infinitesimal model }

### 2023-11-28 review-responses.tex

\reply{
}

\begin{point}{\revref}
In the same paragraph: I don't think the claim that ``independence of offsprings' deviations and midparent values implies a Gaussian distribution'' (which is also stated in the abstract) is true? Surely, ``independence of offsprings' deviations and midparent values'' is just the basic statement of the infinitesimal model (say, in Barton et al 2017) and the infinitesimal model does not imply a Gaussian distribution of trait values? See also section 3.
\end{point}

\reply{
}

\begin{point}{}
    Section 2 and Figure \ref{fig:example_snps}: It is not obvious to me why one would study the ``frequency-weighted'' effect sizes. There is an assumption here I think that allele frequencies are uncorrelated with effect sizes: this will not be true if for example traits are under some kind of stabilising selection: see e.g., \citet{simons2018population} for a detailed exploration of this. Moreover, if SNPs are not truly causal but only tag causal alleles, then weighting by allele frequencies can be even more problematic?

@@

As an alternative, would it make sense to look at what fraction of loci have effect size difference between alternative alleles to be $>2t$ without weighing by allele frequency?
More broadly, how do we interpret the frequency-weighted vs. unweighted distributions?
\end{point}

\reply{
    Good question; this was somewhat obscure in the previous draft.
    We've added a comment which hopefully explains why the frequency-weighted distribution
    is the correct thing to consider for this calculation \revref,
    and have added some more discussion about the point in general
    to the Discussion \llname{discussion_DFEs}.
}

@@

\begin{point}{}
I was also wondering about the ``goodness of fit'' of observed distributions to power laws. I am not necessarily recommending a detailed quantitative analysis, but nevertheless it may be useful to comment on whether one has power to say reject alternative distributions (e.g., a mixture of exponentials) that are often considered. If not, then perhaps the exact exponents inferred are unimportant, and what the data points to more broadly, is a broad distribution of effect sizes.
\end{point}

\reply{
}

\begin{point}{}
    Section 3, Sentence starting ``We know from \citet{barton2017infinitesimal}, \ldots'' \revref: As far as I understand, the statement of \citet{barton2017infinitesimal} is that the joint distribution of trait values of offspring is multivaraite normal, conditioned on the two parents. I don't think the joint distribution of two randomly picked parents and their offspring is mutivariate normal, as this would require some kind of normality assumption about trait values in the population as a whole. It'd be helpful to distinguish more carefully between conditioned and unconditioned distributions, especially in this section.
\end{point}

\reply{
    Looking back at \citet{barton2017infinitesimal}, we see where the confusion comes from,
    and have added some clarifying remarks at the end of this paragraph \revref.
    % Quoting \citet{barton2017infinitesimal},
    % ``Even though the trait distribution across the whole population may be far from Gaussian,
    % the multivariate normal will play a central r\hat{o}le in our analysis''.
    In more detail:
    \citet{barton2017infinitesimal} (``main result'', equation 9)
    shows that conditioned on the trait values in generation $t$,
    the values of generation $t+1$ are also multivariate Gaussian.
    This implies that the joint distribution of traits of any subset of individuals
    is multivariate Gaussian (possibly conditioned on the initial generation)
    -- as \citet{barton2017infinitesimal} says,
    \textit{``equally we could have conditioned on the values of any subset of relatives
    and the same would hold true: expected trait values would be
    a linear functional (determined by the pedigree) of the values
    on which we conditioned, but segregation variances would be unchanged.''}
    (More generally, if $X$ is multivariate Gaussian, and $(Y|X)$ is also multivariate Gaussian
    with mean but not covariance depending on $X$,
    then $(X,Y)$ are jointly multivariate Gaussian.)
    We now also discuss the point further in a subsequent paragraph \llname{pop_not_gaussian}.
}

@@

\begin{point}{}
    In general, I find this section rather confused: the infinitesimal model basically implies that the ``Mendelian segregation term'' $R_{M}$ (which is the difference between an individual and the midparent value $\bar{Z}_{M}$) is normally distributed with mean zero and variance that depends on the relatedness between parents but is independent of the midparent value $\bar{Z}_{M}$. However, in Proposition \ref{prop:parental_contribs}, the authors claim that $\bar{Z}_{M}$ and $R_{M}$ are jointly Gaussian (which is equivalent to saying that the trait value of an individual and the mean of its parents' trait values is jointly Gaussian), which is in general not true.
\end{point}

\reply{
    If the reviewer had said ``confusing'' rather than ``confused'',
    then we would heartily agree.
    After some difficult thinking through of things,
    we've provided a good bit of additional clarification
    about what the Proposition means and doesn't mean:
    see in particular \llname{pop_not_gaussian} and \revref.
}

@@

\begin{point}{}
    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref. As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
\end{point}

\reply{
    Thanks -- we've changed this to ``since models can be easily set up for which the distribution
    of Mendelian sampling terms is not Gaussian, \ldots''
    and provided additional context \revref.
}

@@

\begin{quote}
In Section 5, it is pointed out why the Barton et al. (2017) fails in the case of ‘Cauchy
noise’ (and presumably for any other limit law with $\alpha < 2$). In the absence of giving the
estimate in App D in Barton et al, the explanation at the bottom of p 12 is difficult to
follow. Please provide their estimate for easier comparison (do not assume that readers
studied Barton et al; assume that they scanned (parts) of it).
\end{quote}

\begin{quote}
Finally, Section 6 presents an insightful comparison of the distributions of trait values
resulting from either Gaussian or Cauchy mutation-effect distributions. This seems to be
an explicit genetic model, but again, what is the mutation rate? Clearly, the mutation
rate determines the variance (or more generally variability) of the population distribution.
Please clarify the model. It also needs to be pointed out clearly that such results seem to
be at variance with all (?) explicit multilocus models that have been studied so far

In summary, this ms contains several interesting aspects and will deserve publication after
improving the presentation and clarifying a number of issues.
\end{quote}

\begin{point}{}
Intro: Not too many readers will be familiar with $\alpha$-stable distributions. It may be
useful to clearly point out the difference to the Gaussian case and what happens if $\alpha \ge 2$
(obviously, this is of relevance).
\end{point}

\reply{
    Good idea; we've added this \revref.
}

\begin{point}{}
p 3: State that M is the number of loci.
\end{point}

\reply{
    Whoops - done \revref.
}

@@

\begin{point}{\revref}
Should this read ‘between 20\% and 10\%’? Why this choice?
\end{point}

\reply{
    No, but as written it was confusing, and we've clarified \revref.
    We chose these values by empirical inspection:
    the lower bound was 20 SNPs so that it was not unduly affected by a few large SNPs,
    and the upper bound as 10\% because we wanted the ``tail'', not the ``bulk''.
    (We also checked that changing these values somewhat didn't affect results.)
}

@@

\begin{point}{\revref}
First line in Sect 3: add ‘as $M \to \infty$’. In the next line, this is first time that $\alpha = 2$ is
mentioned. Clarify earlier that this case is very different from $\alpha < 2$ (maybe already below
the last display eq on p3).
\end{point}

\reply{
    Good idea -- done \revref. (Also see \revreffull{2}{1}.)
}

\begin{point}{\revref}
What means ‘nearby alleles’ ? Do you mean physically linked
    loci? Another \llname{loci_polymorphic}: alleles are not polymorphic; loci can be polymorphic.
\end{point}

\reply{
    Good points; both are fixed: \revref and \llname{loci_polymorphic}.
}

@@

\reply{
}

\begin{point}{}
p 12, last display equation: Please relate this explicitly to the corresponding result in
Barton et al. (e.g., by stating their estimate).
\end{point}

\reply{
}

\begin{point}{\revref}
This discussion is too vague. At least, adding a summary of the main findings
would be appropriate (always noting for which version of the infinitesimal model).
\end{point}

\reply{
    We have added a good bit more discussion, including a summary \revref.
}

\begin{point}{\revref}
Does ‘Lévy stable distribution’ refer to the family of $\alpha$-stable
distributions? (I don’t think it has not been used before.)
\end{point}

\reply{
    That's right; changed to ``$\alpha$-stable distribution'' \revref.
}

### 2023-11-28 writeup.tex
Update writeup.tex
many responses

\comment{Make the point that another way to think about distibutions with non-finite variance
    is that there's only a finite number of alleles above any fixed value;
    and also the variance in any finite population will be finite.
    (see reviewer comment, ``all the empirical trait distributions they find have mean and variance'')
    Also discuss comparison to fat-tailed dispersal:
    the question is, what distribution gives us a better model?
    For dispersal, a fat-tailed distribution has in practice finite variance
    but is a better model in that it describes noisiness better;
    here, the analogy is something about polygenicity and noisiness of the process of adaptation.
    If we're modeling non-polygenic (oligogenic) traits then
    why stable as opposed to some other arbitrary distribution?
    Well, the distribution of per-allele effect sizes might not be stable, but CLT tells us that maybe
    stable gives a good approx to the process (as in Gaussian case).
}

@@

As suggested by its name,
under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many - if not most - lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps the distribution of effect sizes within each gene is Normal, \revpoint{1}{13}
but with a standard deviation that varies with ``proximity'' to the trait;
such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus.
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.

%%%%%%%%%%%%%%%%%%%%%%
\section{Empirical observations}

@@

Since the mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta| = |\eta|/2$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen allele at a uniformly chosen locus.
So, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail
(which we take to be all values of $t$ for which the number of SNPs with absolute effect above $t$
is between 20 and 10\% of all SNPs). \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$,
then with $\alpha=2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of the Mendelian sampling term implies a Gaussian distribution.
To see this,
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that 
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

% A similar statement holds for the offspring of two parents in a pedigree,
% using the Skitovich-Darmois theorem \citet{kagan1973characterization},
% except that the portion of the genome shared in common by the parents is unconstrained.

Since models can be easily set up for which trait distributions are not Gaussian,
the implication of this is that for such models, independence of the Mendelian sampling term
is not likely a good assumption. \revpoint{1}{7}
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between nearby allelic effects. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.

As we discuss more later,
the presence of a large effect polymorphic allele
makes $R$ and $\bar Z$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

@@

%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

@@

Finally, we note that above we've most often used the Cauchy distribution
not because it seems most realistic
(indeed data suggests the L\'evy stable distribution with $\alpha=3/2$ might be better)
but for mathematical convenience.

>>>>

\comment{ADD TO DISCUSSION, WHICH IS SHORT:
    Make the point that another way to think about distibutions with non-finite variance
    is that there's only a finite number of alleles above any fixed value;
    and also the variance in any finite population will be finite.
    (see reviewer comment, ``all the empirical trait distributions they find have mean and variance'')
    Also discuss comparison to fat-tailed dispersal:
    the question is, what distribution gives us a better model?
    For dispersal, a fat-tailed distribution has in practice finite variance
    but is a better model in that it describes noisiness better;
    here, the analogy is something about polygenicity and noisiness of the process of adaptation.
    If we're modeling non-polygenic (oligogenic) traits then
    why stable as opposed to some other arbitrary distribution?
    Well, the distribution of per-allele effect sizes might not be stable, but CLT tells us that maybe
    stable gives a good approx to the process (as in Gaussian case).
}

@@

As suggested by its name,
under the infinitesimal model the effects of any given genetic variant on the trait
is small (indeed, vanishing),
and selection-driven trait shifts are mediated by small changes in frequencies of many alleles of small effect.
However, empirical understanding of the genetic basis of trait variation in real-world species
often finds a significant portion of heritable variation explained by variants at only a few loci.
There are many possible explanations for this observation,
including selection or publication bias,
filters for larger-effect variants due to spatially varying selection
\citep{barton1987establishment,yeaman2011establishment,yeaman2011genetic},
and/or grouping together of concordant alleles by linkage or inversions
\citep{kirkpatrick2006chromosome}.
However,
it seems clear that in at least those cases where we understand the underlying molecular mechanisms,
mutations of large effect can be important in practice.
% \citep[e.g.,][]{cyclodiene-resistance,hoekstra2006genetics}.
The consensus in the field seems to be that traits exist on a spectrum between monogenic and polygenic,
and that many -- if not most -- lie in the intermediate region,
where heritable trait variation is due both to segregating large- and small-effect variants.
Indeed, some sensible verbal models of the biological mechanisms by which a genetic change
comes to have an effect on an organism's trait
describe a hierarchy or range of ``proximities'' to the trait,
by which changes to more ``nearby'' genes have the possibility of causing larger changes in the trait
\citep{kopp2012transcriptional,boyle2017expanded}.
This suggests a mixture model for the effect size distribution:
perhaps the distribution of effect sizes within each gene is Normal, \revpoint{1}{13}
but with a standard deviation that varies with ``proximity'' to the trait;
such mixtures can easily have large proportions of extreme values.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt(\log n)$. \revpoint{2}{1}

@@

Suppose that the focal individual's trait is
$$
    Z_M = \sum_{\ell=1}^M \eta_\ell / M^{1/\alpha},
$$
where $\eta_\ell$ is the contribution of the allele carried at the $\ell^\text{th}$ locus,
and $M$ is the number of loci. \revpoint{2}{2}
The effects of the alleles carried by the two parents
are $\eta_\ell^{[1]}$ and $\eta_\ell^{[2]}$, respectively,
and their traits are
$Z_M^{[1]} = \sum_\ell \eta_\ell^{[1]} / M^{1/\alpha}$ and
$Z_M^{[2]} = \sum_\ell \eta_\ell^{[2]} / M^{1/\alpha}$.
We assume the loci are unlinked and inheritance is fair, so
letting $X_\ell$ be a family of Bernoulli(1/2) random variables that are independent of everything else,
$$
    \eta_\ell = X_\ell \eta_\ell^{[1]} + (1 - X_\ell) \eta_\ell^{[2]} .
$$
What we will call the \emph{Mendelian sampling term}
(or occasionally \emph{Mendelian noise}), $R_M$,
is the offspring's trait minus the midparent value, $\bar Z_M$:
\begin{align*}
    \bar Z_M &= \frac{Z_M^{[1]} + Z_M^{[2]}}{2} \\
    R_M &= Z_M - \bar Z_M \\
        &= \sum_{\ell=1}^M
            \frac{
                X_\ell (\eta_\ell^{[1]} - \eta_\ell^{[2]})
                + (1 - X_\ell) (\eta_\ell^{[2]} - \eta_\ell^{[1]})
            }{ 2 } .
\end{align*}
Note in particular that
the contribution of locus $\ell$ to $R_M$ is $\pm |\eta_\ell^{[1]} - \eta_\ell^{[2]}|$.

%%%%%%%%%%%%%%%%%%%%%%
\section{Empirical observations}
    \label{sec:empirical}

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
This will make a difference as frequencies and effect sizes are often correlated in practice.
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

We know from \citet{barton2017infinitesimal} \revpoint{1}{5} \revpoint{2}{4}
that if the $\eta^{[i]}_\ell$ are
independent with mean zero and $\sum_{\ell=1}^M \var[\eta_\ell]/M \to \sigma^2$ as $M \to \infty$,
then with $\alpha \ge 2$,
the traits $(Z_M, Z_M^{[1]}, Z_M^{[2]})$ converge in distribution to a multivariate Gaussian distribution
$(Z, Z^{[1]}, Z^{[2]})$ for which $Z - (Z^{[1]} + Z^{[2]})/2$ is independent of $Z^{[1]}$ and $Z^{[2]}$.
(The main result in \citet{barton2017infinitesimal} is that the trait values in generation $t$
are Gaussian conditioned on the traits in generation $t-1$,
but as discussed in that paper,
this implies that the traits of any set of individuals is multivariate Gaussian,
with covariance matrix determined by the pedigree.
Note, however, that this does \emph{not} imply that the distribution of trait values
across the population is Gaussian, as a collection of correlated Gaussians is not itself Gaussian.)

Perhaps surprisingly, there is a converse --
in at least some situations,
independence of the Mendelian sampling term \emph{implies} a Gaussian distribution.
To see this (and to see exactly what we mean by this),
divide the parental loci into groups according to the results of inheritance,
defining
\begin{align*}
    U_M^{[1]} &= \sum_{\ell=1}^M X_\ell \eta_\ell^{[1]}  ,
    \qquad \qquad
    V_M^{[1]} = \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[1]} ,  \\
    U_M^{[2]} &= \sum_{\ell=1}^M (1-X_\ell) \eta_\ell^{[2]} ,
    \qquad \qquad
    V_M^{[2]} = \sum_{\ell=1}^M X_\ell \eta_\ell^{[2]}  ,
\end{align*}
so that 
$Z_M^{[1]} = U_M^{[1]} + V_M^{[1]}$ and
$Z_M^{[2]} = U_M^{[2]} + V_M^{[2]}$, while
$Z_M = U_M^{[1]} + U_M^{[2]}$ is the transmitted portion of the parental genomes,
and we can define $V_M = V_M^{[1]} + V_M^{[2]}$ to be the untransmitted portion.
Furthermore,
\begin{align*}
    \bar Z_M = \frac{Z_M + V_M}{2}
    \qquad \text{ and } \qquad
    R_M = \frac{Z_M - V_M}{ 2 } .
\end{align*}

@@

We should now discuss what this counter-intuitive result says and does not say.
The proposition says, roughly, that if the parental alleles are unlinked and independently chosen,
and meotic segregation is fair,
then independence of Mendelian sampling term from midparent value
\emph{implies} a joint Gaussian distribution for the two.
We have not assumed anything about the distribution of the trait in the population:
rather, we have showed that any model of this sort
that assumes independence of Mendelian sampling term and midparent value
must also, by logical necessity, have a Gaussian distribution for the two.
In other words,
since models can be easily set up for which the distribution of Mendelian sampling terms
is not Gaussian,
the implication of the proposition
is that for such models, independence of the Mendelian sampling term
is not likely a good assumption. \revpoint{1}{7}
Note that the proof does not depend on independence of the $X_\ell$:
indeed, alleles may be inherited together,
although in practice this would quickly create dependencies between the effects
of alleles at such linked loci. \revpoint{2}{5}
Also note that it is \emph{not} usually true that the components --
for instance, $U_M^{[1]}$ and $U_M^{[2]}$ --
are independent of each other.
Finally, note that although they may not be independent,
the Mendelian sampling term $R_M$ and the midparent value $\bar Z_M$
\emph{are} always uncorrelated, since $Z_M$ and $V_M$ are exchangeable.

Note that under the conditions of the proposition, \llabel{pop_not_gaussian}
the parental traits, $Z^{[1]}_M$ and $Z^{[2]}_M$,
are also (jointly) Gaussian.
(This was not an assumption, however.)
If parents are chosen uniformly, then
the distribution of, say, $Z^{[1]}$ is just the population's trait distribution.
This seems wrong, since it is well-known that forces such as selection or migration
can make the population distribution non-Gaussian.
However, there is no contradiction:
we have stated the proposition only in terms of the \emph{first} generation
of offspring of unrelated parents \citep[as in][]{barton2017infinitesimal}.
Proposition~\ref{prop:parental_contribs} does not imply that
the distribution of traits in the offspring generation is Gaussian:
unless all offspring are unrelated,
it will be a collection of correlated Gaussians,
which is not itself Gaussian.
(As an extreme example, suppose that half the offspring are produced by the two parents
with largest trait values, and the other half by the parents with the smallest trait values;
then although each offspring's distribution about the midparent is Gaussian,
the offspring's population distribution is a mixture of two Gaussians
about these two midparents.)
What about further generations?
A similar statement to the proposition
holds for the offspring of two parents in an arbitrary pedigree,
using the Skitovich-Darmois theorem \citet{kagan1973characterization}
% except that the portion of the genome shared in common by the parents is unconstrained.
-- however, there is nothing to be gained by the additional generality.

As we discuss more later,
the presence of a large effect allele at a polymorphic locus \llabel{loci_polymorphic}
makes $R$ and $\bar Z$ dependent:
consider the extreme case where $\eta^{[i]}_\ell$ are i.i.d.{} with variance 1 for $\ell \ge 2$,
while $\eta^{[i]}_1 = \pm 10$ with probability 1/2 each,
and $Z^{[i]} = \eta_1^{[i]} + \sum_{\ell=2}^M \eta_\ell^{[i]} / \sqrt{M}$.
Then in the limit $R$ depends strongly on parental trait values:
if $Z^{[1]} \approx Z^{[2]}$ and so $\bar Z \approx \pm 10$ then most likely
$\eta_1^{[1]} = \eta_1^{[2]}$ and so $R$ is Normal with variance 1/2;
but if $\eta_1^{[1]} \neq \eta_1^{[2]}$ (and so $\bar Z \approx 0$)
then $R$ is close to $+5$ or $-5$ with equal probability.

Finally, a possible point of confusion is that \revpoint{1}{6}
the infinitesimal model is often phrased as a statement about offspring trait distributions
\emph{given} parental trait values,
while our statement above regards parental trait values as random.
However, the notions are equivalent:
the statement that
``for each pair of parents the deviation of the offspring's trait from midparent
has a distribution that does not depend on the parents' values''
is essentially the definition of statistical independence
of offspring's Mendelian sampling term and parental trait values.

@@

%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
    \revpoint{2}{11}

The infinitesimal model is a powerful and tractable model of trait inheritance,
but is polygenic to a fault,
excluding the possibility of loci of large effect
(that are often seen in practice).
In this paper,
we have looked at the possibility of using a broader range of central limit theorems
to extend the infinitesimal model to traits for which 
effect size distributions fall in the domain of attraction of the $\alpha$-stable distributions
(i.e., a larger proportion of variance
is explained by a smaller number of loci).
Indeed, we found evidence that this is true of many real traits,
although our relatively uncritical use of GWAS estimates makes this evidence suggestive at best.
Next, we show that in fact the infinitesimal model \emph{cannot} describe non-Gaussian distributions
\comment{XXX TODO}.
A recurring theme in the rest of the paper is
that just as the infinitesimal model is useful precisely because it ignores the genetic details,
it cannot apply once some genetic details are important
(i.e., there are loci of large effect).

@@

Finally, we note that above we've most often used the Cauchy distribution
not because it seems most realistic
(indeed data suggests the $\alpha$-stable distribution with $\alpha=3/2$ might be better) \revpoint{2}{12}
but for mathematical convenience.

\subsection*{Selection and effect size distributions} \llabel{discussion_DFEs}
A substantial body of theoretical work has used various models and assumptions
to predict what distributions of effect sizes are expected based on biology.
In comparing to this work, there are several distinctions that are important to make.
For instance, selection makes the distribution of effect sizes identified by GWAS
different than the distribution of effect sizes of \emph{de novo} mutations,
and predictions are made about this relationship by \citet{simons2018population} % and \citet{simons2022simple}.
(under stabilizing selection)
and \citet{hayward2022polygenic}
(under adaptation to a new environment).
Our first question in this paper is orthogonal to this discussion:
\emph{given} the available genetic variation,
what model should we use to predict the short-term course of evolution?
As we saw in Section~\ref{sec:empirical},
the frequency-weighted effect size distribution (perhaps closer to that seen by GWAS)
is more relevant for this question than the \emph{de novo} effect size distribution.
In the latter part of this paper we look at long-term dynamics,
but with major differences:
most obviously, we only consider the distribution of the Mendelian noise term,
rather than the distribution of individual allelic effects.
These two are clearly related, but a more detailed analysis is beyond the scope of this paper.
